Individual Variation in Animal Movement and Effects on Seed Dispersal Distance

### 2021-01-04 paper/paper.Rmd
Updated introduction

---
title: "Individual Variation in Animal Movement and Effects on Seed Dispersal Distance"
abstract: |
  Seed dispersal is a fundamental process in plant ecology and due to its high importance in the ecosystem, various modelling approaches have been developed to understand seed movement and dispersal mechanisms. In particular, long-distance seed dispersal (LDD) plays a significant role in determining a population's genetic diversity and range expansion. Animal-mediated seed dispersal is a key component of LDD events, as frugivores can travel long distances and transport seeds with them. In this study we seek to understand the implications of individual variation in animal movement and how these can impact estimates of long distance seed dispersal. We apply these methods to empirical data sets that contain locations for an araçari, *Pteroglossus pluricintus*, one of the primary frugivores for the Amazonian canopy tree *Virola flexuosa*. We combine animal movement data and gut retention time to simulate *Virola* seed dispersal. This approach showed that there is significant variation in individual araçari movement patterns which directly influence the estimated number of long-distance dispersal events for *Virola* seeds. We found that models that ignore this underlying variation in individual animal movement, underestimate the percentage of long-distance seed dispersal events and maximum seed dispersal distances.
output:
  word_document: default
  html_document:
    df_print: paged
bibliography: mybibfile.bib
---

@@

Differences in behavior among individuals (i.e. behavioral syndromes sensu [@sih2004behavioral]) can have remarkable consequences on ecological and evolutionary processes [@wolf_animal_2012]. For vertebrate dispersed plants, animal behavior will inﬂuence seed shadows and few species, or even individuals, might provide unique, high value dispersal services by moving farther distances from parental plants [@schupp_seed_2010]. Long distance dispersal (LDD) events are crucial to reaching suitable sites to germinate and establish and to colonize new habitats [@nathan2006long]. Furthermore, rare events (often overlooked) are critical for dispersal and might have a large effect on resulting plant demography [@loayza2014seed]. Incorporating animal behavior and the individual distances they carry seeds into seed dispersal models might result in better predictions of seed shadow and better inform about the resulting patterns of seed dispersal. To understand the consequences of variation in animal movement, we compare seed dispersal estimates from simulation models that incorporate seed retention time and individual, family, or population level animal movement.

>>>>

---
title: "Individual Variation in Animal Movement and Effects on Seed Dispersal Distance"
abstract: |
  I don't like this abstract. Revisit at the end...Seed dispersal is a fundamental process in plant ecology and due to its high importance in the ecosystem, various modelling approaches have been developed to understand seed movement and dispersal mechanisms. In particular, long-distance seed dispersal (LDD) plays a significant role in determining a population's genetic diversity and range expansion. Animal-mediated seed dispersal is a key component of LDD events, as frugivores can travel long distances and transport seeds with them. In this study we seek to understand the implications of individual variation in animal movement and how these can impact estimates of long distance seed dispersal. We apply these methods to empirical data sets that contain locations for an araçari, *Pteroglossus pluricintus*, one of the primary frugivores for the Amazonian canopy tree *Virola flexuosa*. We combine animal movement data and gut retention time to simulate *Virola* seed dispersal. This approach showed that there is significant variation in individual araçari movement patterns which directly influence the estimated number of long-distance dispersal events for *Virola* seeds. We found that models that ignore this underlying variation in individual animal movement, underestimate the percentage of long-distance seed dispersal events and maximum seed dispersal distances.
output:
  word_document:
    reference_docx: paper_ref.docx
  html_document:
    df_print: paged
bibliography: mybibfile.bib
---

@@

In this paper, we focus on the implications of intraspecific variation in animal movement for seed dispersal distances and seed aggregation across the landscape. To understand the consequences of this variation in movement, we developed a spatially-explicit individual-based model of the many-banded aracari, *Pteroglossus pluricinctus*, one of the primary frugivores of the Amazonian canopy tree *Virola flexuosa*. Also, previous studies focused on this system have shown that *P. pluricintus* is able to disperse seeds across long distances [@holbrook_using_2007; @holbrook_dispersal_2009; @holbrook_home_2011], an important for plant population. Long distance dispersal (LDD) events are crucial to reaching suitable sites to germinate and establish and to colonize new habitats [@nathan2006long]. Furthermore, rare events (often overlooked) are critical for dispersal and might have a large effect on resulting plant demography [@loayza2014seed]. Our first objective in this work was to study the differences of simulated seed shadows between models with and without underlying variability in animal movement. The second objective focused on quantifying the differences of LDD events reflected between seed shadows. Lastly, our third objective focused on the spatial spread between seeds, and how individual variation in animal movement affects seed aggregation. Identifying the implications of individual variation in animal movement over seed dispersal estimates can help us improve future models and understand the effects of frugivore traits on plant population dynamics.
